Galactic Chemical Evolution Simulations Favor Long Delay Times for Type Ia Supernovae

### 2023-12-11 src/tex/ms.tex
[showyourwork] overleaf sync

Figure \ref{fig:ofe-feh-dtd} compares [O/Fe]--[Fe/H] distributions for five simulations with the same SFH (inside-out) but different DTD models. We choose the inside-out SFH model for this figure because it produces a monotonically-decreasing abundance track, making comparisons between the different DTD models relatively straightforward. We note that in this model, both the low- and high-$\alpha$ sequences are composed of stars with a wide range of birth $R_{\rm gal}$. The simulation results are arranged so that from left to right the fraction of prompt SNe Ia decreases and the mean delay time increases. Results for the two-population and power-law ($\alpha$=-1.1) DTD models, with a large fraction of prompt SNe Ia, are presented in the first and second columns, respectively. For these models, the simulated stellar abundances are reasonably well-aligned with the APOGEE contours at low $|z|$ (bottom panels), but they entirely miss the observed high-$\alpha$ sequence at high $|z|$ (top panels). In the center column, the exponential ($\tau=1.5$ Gyr) DTD produces a distribution which aligns quite well with the APOGEE contours in all $|z|$-bins, and even produces a ``ridge'' which extends to large [O/Fe] at low- and mid-latitudes (bottom and center panels, respectively). While it does better at populating the high-$\alpha$ sequence than the previous DTD models, the bulk of the simulated populations at large $|z|$ still fall below the APOGEE 30\% contour (top panel). The two right-hand columns present simulation results for the plateau ($W=1$ Gyr) and triple-system DTD models,

@@

>>>>

Figure \ref{fig:ofe-feh-dtd} compares [O/Fe]--[Fe/H] distributions for five simulations with the same SFH (inside-out) but different DTD models. We choose the inside-out SFH model for this figure because it produces a monotonically-decreasing abundance track, making comparisons between the different DTD models relatively straightforward. We note that in this model, both the low- and high-$\alpha$ sequences are composed of stars with a wide range of birth $R_{\rm gal}$. The simulation results are arranged by increasing median delay time from left to right. Results for the two-population and power-law ($\alpha$=-1.1) DTD models, with a large fraction of prompt SNe Ia, are presented in the first and second columns, respectively. For these models, the simulated stellar abundances are reasonably well-aligned with the APOGEE contours at low $|z|$ (bottom panels), but they entirely miss the observed high-$\alpha$ sequence at high $|z|$ (top panels). In the center column, the exponential ($\tau=1.5$ Gyr) DTD produces a distribution which aligns quite well with the APOGEE contours in all $|z|$-bins, and even produces a ``ridge'' which extends to large [O/Fe] at low- and mid-latitudes (bottom and center panels, respectively). While it does better at populating the high-$\alpha$ sequence than the previous DTD models, the bulk of the simulated populations at large $|z|$ still fall below the APOGEE 30\% contour (top panel). The two right-hand columns present simulation results for the plateau ($W=1$ Gyr) and triple-system DTD models, which have the longest median delay times. The high-$\alpha$ ``knee'' occurs at a much higher metallicity in these models and is visible in the gas abundance tracks in the upper-left corner of the panels. At high $|z|$, the simulated abundance align quite well with the APOGEE high-$\alpha$ sequence (top panels), but there is a significant ridge of high-$\alpha$ stars from the inner Galaxy at low $|z|$ (bottom panels).

@@

Figure \ref{fig:age-ofe-sfh} compares the stellar age and [O/Fe] distributions in the Solar neighborhood between our four SFH models. Similar to Figure \ref{fig:ofe-feh-sfh}, all four panels assume an exponential DTD model with $\tau=1.5$ Gyr. We include age data from \citet{Leung2023-Ages} for qualitative comparison, though we caution against drawing strong conclusions because we don't correct for selection effects or systematics in the neural net ages. The inside-out and late-burst models show fair agreement with the data at high [O/Fe], although both show a $\sim2$ Gyr offset to older ages. One could shift the ramp-up in star formation to slightly later times or simply run the simulation for a shorter amount of time to close this gap. Like the data, the trend in the median age with decreasing [O/Fe] decreases monotonically in the inside-out model. The late-burst model, however, shows a bump in the distribution around a lookback time of $\sim2$ Gyr driven by the Gaussian SFR burst, which is not seen in the data, as noted by \citet{Johnson2021-Migration}. Neither model produces a significant amount of stars below Solar [O/Fe], in contrast to the data.

For the early-burst SFH model, the simulated stellar ages are almost perfectly aligned with the data for [O/Fe]$\gtrsim 0.2$. The rapid rise in the SFE at early times delays the descent to lower [O/Fe] values and produces a clump of low-metallicity, high-[O/Fe] stars at an age of $\sim10$ Gyr. The early-burst model produces a sharp transition between a linear relationship between [O/Fe] and age for stars older than $\sim5$ Gyr and a flat relationship for the younger stars, and it too misses the lowest [O/Fe] bin in the data. Finally, the two-infall SFH model produces the closest overall agreement between the simulated and observed median ages. Stars with [O/Fe]$\gtrsim 0.25$ were produced in the first infall, while the second infall produces a clump of stars with similar metallicity, ages of $\sim8$ Gyr, and [O/Fe]$\approx0.2$. There is an interesting small population of very old, low-$\alpha$ stars which is not seen in any of the other models. The two-infall model is the only one which plateaus at sub-Solar [O/Fe] for the youngest stars, leading to a very good match to the observed data at low [O/Fe].
